---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

```{r}
library(ggplot2); library(MASS); library(splines); library(rstan)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
```

Load virus data
```{r}
covid <- read.csv("/Users/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/04-20-2020.csv")
# covid <- read.csv("/Volumes/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/04-20-2020.csv")
#Subset to get only US
covid <- covid[covid$Country_Region == "US" ,]
#Remove non-state entries
covid <- covid[nchar(as.character(covid$Admin2)) > 0, ]
#Get state abbreviations in from a key file
state.abb <- read.csv("/Users/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/state_abb_key.csv")
# state.abb <- read.csv("/Volumes/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/state_abb_key.csv")
#Make a new column with state abbreviations
state.abb$State <- as.character(state.abb$State)
state.abb$State <- gsub('^\\s', '', state.abb$State)
state.abb$Abb <- as.character(state.abb$Abb)
covid$Province_State <- as.character(covid$Province_State)
#Loop through and grab the abbreviation for each state
covid$state_abb <- 1:nrow(covid)
for (i in 1:nrow(covid)) {
  covid$state_abb[i] <- (state.abb$Abb[state.abb$State == covid$Province_State[i]])
}
#Paste with county to make state_county column
covid$state_county <- paste(covid$state_abb, covid$Admin2, sep = "_")
```

Load population data by county
```{r}
pop <- read.csv("/Users/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/co-est2019-alldata.csv")
# pop <- read.csv("/Volumes/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/co-est2019-alldata.csv")
#Turn 'Doña Ana County' in NM to 'Dona Ana'
pop$CTYNAME <- as.character(pop$CTYNAME)
pop$CTYNAME[1835] <- "Dona Ana County"
#Subset to get rid of any row without a comma (full states)
pop <- pop[!(pop$CTYNAME %in% grep(' [CountyParish]', as.character(pop$CTYNAME), invert=T, value=T)), ]
# pop$CTYNAME <- as.factor(pop$CTYNAME)
#Strip "County" from the countyname and make a State_County column in the population dataset
# pop$county <- gsub(" County", "", pop$CTYNAME)
#Make county column
pop$county <- gsub('\\sParish', '', gsub('\\sCounty', '', pop$CTYNAME))
#Get state abbreviations in from a key file
state.abb <- read.csv("/Users/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/state_abb_key.csv")
# state.abb <- read.csv("/Volumes/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/state_abb_key.csv")
#Make a new column with state abbreviations
state.abb$State <- as.character(state.abb$State)
state.abb$State <- gsub('^\\s', '', state.abb$State)
state.abb$Abb <- as.character(state.abb$Abb)
pop$STNAME <- as.character(pop$STNAME)
#Loop through and grab the abbreviation for each state
pop$state_abb <- 1:nrow(pop)
for (i in 1:nrow(pop)) {
  pop$state_abb[i] <- (state.abb$Abb[state.abb$State == pop$STNAME[i]])
}
#Paste with county to make state_county column
pop$state_county <- paste(pop$state_abb, pop$county, sep = "_")
```

Load land area data by county
```{r}
area <- read.csv("/Users/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/LND01.csv")
# area <- read.csv("/Volumes/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/LND01.csv")
#Subset to get rid of any row without a comman (full states and the US)
area <- area[!(area$Areaname %in% grep('\\,', area$Areaname, invert = T, value = T)), ]
#Strip away county info to get just state
area$state <- gsub("[A-Za-z \\t\\'\\.\\-]+\\,\\s", "", area$Areaname)
#Strip away state info to get county
area$county <- gsub("\\, [A-Z]+", "", area$Areaname)
#Combine them to get State_County that matches the pop dataframe
area$state_county <- paste(area$state, area$county, sep="_")
```

Healthcare coverage
```{r}
coverage <- read.csv("/Users/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/SAHIE_21APR20_11_22_31_15.csv")
# coverage <- read.csv("/Volumes/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/SAHIE_21APR20_11_22_31_15.csv")
#Remove whole-state entries
coverage <- coverage[coverage$Name %in% grep('\\,', coverage$Name, invert = F, value = T), ]
#Make county column
coverage$county <- gsub('\\sParish\\,\\s[A-Za-z\\s \\.]+', '', gsub('\\sCounty\\,\\s[A-Za-z\\s \\.]+', '', coverage$Name))
#For 'counties' that have something like city in them instead of "County"
# coverage$county <- gsub('\\,\\s[A-Z]+', '', coverage$county)
#Make state abb column
coverage$state <- gsub("[A-Za-z \\,\\.\\s\\-\\'\\-]+\\s[PCc][a-z]+\\,\\s", "", coverage$Name)
# coverage$state <- gsub("[A-Za-z \\,\\.\\s\\-\\']+\\sParish\\,\\s", "", coverage$state)
#Remove Alaska, Virginia b/c weird
coverage <- coverage[coverage$Name %in% grep('\\, AK', coverage$Name, invert = T, value = T), ]
# coverage <- coverage[coverage$Name %in% grep('\\, VA', coverage$Name, invert = T, value = T), ]
#Change specific ones to be consistent
coverage$county[coverage$county == "Baltimore city"] <- "Baltimore City"
coverage$county[coverage$county == "St. Louis city"] <- "St. Louis City"
#Make state_county column
coverage$state_county <- paste(coverage$state, coverage$county, sep="_")
```

Hospitals
```{r}
hosp <- read.csv("/Users/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/Hospitals.csv")
# hosp <- read.csv("/Volumes/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/Hospitals.csv")
#Subset only 'General medical and surgical' hospitals
hosp <- hosp[hosp$NAICS_DESC %in% c("CHILDREN'S HOSPITALS, GENERAL", "GENERAL MEDICAL AND SURGICAL HOSPITALS"), ]
#Get rid of AK and VA
hosp <- hosp[!(as.character(hosp$STATE) %in% c("AK")), ]
#Make a lowercase county column
hosp$county <- tolower(hosp$COUNTY)
#Turn any letter at the start of the line or after a space into uppercase
hosp$county <- gsub('^([a-z])', '\\U\\1', gsub('\\s([a-z])', ' \\U\\1', hosp$county, perl=T), perl = T)
#Make a state_county
hosp$state_county <- paste(hosp$STATE, hosp$county, sep="_")
#Loop over unique state_county values and count the number of rows (hospitals)
hosp.count <- 1:length(unique(hosp$state_county))
for (i in 1:length(unique(hosp$state_county))) {
  hosp.count[i] <- nrow(hosp[hosp$state_county == unique(hosp$state_county)[i], ])
}
#Put the hospital counts and state_county columns together
hosp.count.df <- data.frame(hosp=hosp.count, state_county=unique(hosp$state_county))
#Only counties with hopitals are listed here (because obviously a hospital that doesn't exist wouldn't be listed), so let's grab the names of the counties that aren't here and make them have 0 hospitals
no.hosp <- data.frame(hosp=rep(0, length(pop$state_county[!(pop$state_county %in% hosp.count.df$state_county)])), state_county=pop$state_county[!(pop$state_county %in% hosp.count.df$state_county)])
hosp.count.df <- rbind(hosp.count.df, no.hosp)
#Order the dataframe
hosp.count.df <- hosp.count.df[order(as.character(hosp.count.df$state_county)), ]
```

Average weekly salary
```{r}
salary <- read.csv("/Users/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/prior_yr_county_high_level/allhlcn18.csv")
# salary <- read.csv("/Volumes/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/prior_yr_county_high_level/allhlcn18.csv")
#Remove non-counties
salary <- salary[nchar(as.character(salary$St.Name)) > 0 ,]
#Remove state-wide characteristics where Cnty = 0
salary <- salary[salary$Cnty > 0, ]
#Grab only the 'total employees' row for each county
salary <- salary[salary$Ownership == "Total Covered", ]

#Make county column
salary$county <- gsub('\\sCity\\,\\s[A-Za-z\\s \\.]+', '', gsub('\\sParish\\,\\s[A-Za-z\\s \\.]+', '', gsub('\\sCounty\\,\\s[A-Za-z\\s \\.]+', '', salary$Area)))
#For 'counties' that have something like city in them instead of "County"
salary$county <- gsub('\\,\\s[A-Z]+', '', salary$county)
#Remove unknown counties
salary <- salary[salary$county %in% grep('Unknown', salary$county, invert = T, value = T), ]
#Remove Alaska b/c weird
salary <- salary[salary$St.Name != "Alaska", ]
#Get state abbreviations in from a key file
state.abb <- read.csv("/Users/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/state_abb_key.csv")
# state.abb <- read.csv("/Volumes/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/state_abb_key.csv")
#Make a new column with state abbreviations
state.abb$State <- as.character(state.abb$State)
state.abb$State <- gsub('^\\s', '', state.abb$State)
state.abb$Abb <- as.character(state.abb$Abb)
#Loop through and grab the abbreviation for each state
salary$state_abb <- 1:nrow(salary)
for (i in 1:nrow(salary)) {
  salary$state_abb[i] <- (state.abb$Abb[state.abb$State == salary$St.Name[i]])
}
#Make state_county column
salary$state_county <- paste(salary$state_abb, salary$county, sep="_")
```

2016 presidential election votes
```{r}
pres <- read.csv("/Users/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/election-context-2018.csv")
# pres <- read.csv("/Volumes/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/election-context-2018.csv")
#Calculate total votes for 2016 presidential election
pres$votes <- pres$trump16 + pres$clinton16 + pres$otherpres16
#Calculate percentage votes for Trump
pres$trump.per <- pres$trump16/pres$votes
#Get state abbreviations in from a key file
state.abb <- read.csv("/Users/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/state_abb_key.csv")
# state.abb <- read.csv("/Volumes/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/state_abb_key.csv")
#Make a new column with state abbreviations
state.abb$State <- as.character(state.abb$State)
state.abb$State <- gsub('^\\s', '', state.abb$State)
state.abb$Abb <- as.character(state.abb$Abb)
#Loop through and grab the abbreviation for each state
pres$state_abb <- 1:nrow(pres)
for (i in 1:nrow(pres)) {
  pres$state_abb[i] <- (state.abb$Abb[state.abb$State == pres$state[i]])
}
#Make state_county column
pres$state_county <- paste(pres$state_abb, pres$county, sep="_")
```

Previous 2 days of cases
```{r}
#Previous day
prev.day <- read.csv("/Users/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/04-19-2020.csv")
# prev.day <- read.csv("/Users/tschappe/Documents/NCState/Classes/ST540/Final\ Project/04-19-2020.csv")
#Subset to get only US
prev.day <- prev.day[prev.day$Country_Region == "US" ,]
#Remove non-state entries
prev.day <- prev.day[nchar(as.character(prev.day$Admin2)) > 0, ]
#Get state abbreviations in from a key file
state.abb <- read.csv("/Users/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/state_abb_key.csv")
# state.abb <- read.csv("/Volumes/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/state_abb_key.csv")
#Make a new column with state abbreviations
state.abb$State <- as.character(state.abb$State)
state.abb$State <- gsub('^\\s', '', state.abb$State)
state.abb$Abb <- as.character(state.abb$Abb)
prev.day$Province_State <- as.character(prev.day$Province_State)
#Loop through and grab the abbreviation for each state
prev.day$state_abb <- 1:nrow(prev.day)
for (i in 1:nrow(prev.day)) {
  prev.day$state_abb[i] <- (state.abb$Abb[state.abb$State == prev.day$Province_State[i]])
}
#Paste with county to make state_county column
prev.day$state_county <- paste(prev.day$state_abb, prev.day$Admin2, sep = "_")

prev.2day <- read.csv("/Users/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/04-18-2020.csv")
# prev.2day <- read.csv("/Users/tschappe/Documents/NCState/Classes/ST540/Final\ Project/04-18-2020.csv")
#Subset to get only US
prev.2day <- prev.2day[prev.2day$Country_Region == "US" ,]
#Remove non-state entries
prev.2day <- prev.2day[nchar(as.character(prev.2day$Admin2)) > 0, ]
#Get state abbreviations in from a key file
state.abb <- read.csv("/Users/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/state_abb_key.csv")
# state.abb <- read.csv("/Volumes/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/state_abb_key.csv")
#Make a new column with state abbreviations
state.abb$State <- as.character(state.abb$State)
state.abb$State <- gsub('^\\s', '', state.abb$State)
state.abb$Abb <- as.character(state.abb$Abb)
prev.2day$Province_State <- as.character(prev.2day$Province_State)
#Loop through and grab the abbreviation for each state
prev.2day$state_abb <- 1:nrow(prev.2day)
for (i in 1:nrow(prev.2day)) {
  prev.2day$state_abb[i] <- (state.abb$Abb[state.abb$State == prev.2day$Province_State[i]])
}
#Paste with county to make state_county column
prev.2day$state_county <- paste(prev.2day$state_abb, prev.2day$Admin2, sep = "_")

#Combine them
prev.day.out <- subset(prev.day, select=c("FIPS", "Confirmed", "Deaths", "state_county"))
prev.2day.out <- subset(prev.2day, select=c("FIPS", "Confirmed", "Deaths", "state_county"))
prev.day.merged <- merge(prev.day.out, prev.2day.out, by="FIPS")
prev.day.merged$prev.ratio <- prev.day.merged$Confirmed.x/prev.day.merged$Confirmed.y
prev.day.merged$prev.diff <- prev.day.merged$Confirmed.x - prev.day.merged$Confirmed.y
prev.day.merged.out <- subset(prev.day.merged, select=c("prev.ratio", "prev.diff", "state_county.x"))
colnames(prev.day.merged.out)[3] <- c("state_county")
```

Tests per person for each state
```{r}
#Load state population data
state.pop <- read.csv("/Users/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/SCPRC-EST2019-18+POP-RES.csv")
state.pop <- state.pop[!(state.pop$NAME %in% c("United States", "Puerto Rico Commonwealth")), ]
#Get state abbreviations in from a key file
state.abb <- read.csv("/Users/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/state_abb_key.csv")
# state.abb <- read.csv("/Volumes/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/state_abb_key.csv")
#Make a new column with state abbreviations
state.abb$State <- as.character(state.abb$State)
state.abb$State <- gsub('^\\s', '', state.abb$State)
state.abb$Abb <- as.character(state.abb$Abb)
state.pop$NAME <- as.character(state.pop$NAME)
#Loop through and grab the abbreviation for each state
state.pop$state_abb <- 1:nrow(state.pop)
for (i in 1:nrow(state.pop)) {
  state.pop$state_abb[i] <- (state.abb$Abb[state.abb$State == state.pop$NAME[i]])
}

#Load COVID testing data
covid.test <- read.csv("/Users/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/covid_testing.csv")
covid.test <- covid.test[covid.test$State %in% state.pop$state_abb, ]
covid.test$tests <- as.integer(covid.test$tests)
pop.temp <- NULL
for (i in 1:nrow(covid.test)) {
  pop.temp[i] <- state.pop$POPESTIMATE2019[as.character(state.pop$state_abb) == as.character(covid.test$State[i])]
}
covid.test$pop <- as.integer(pop.temp)
#Calculate COVID tests per person
covid.test$test.per.pop <- covid.test$tests/covid.test$pop
```

Socioeconomic vulnerability index
```{r}
svi <- read.csv("/Users/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/SVI2018_US_COUNTY.csv")
#Remove 'City' from county column
svi$county <- gsub('\\sCity', '', svi$COUNTY)
#Make state_county column
svi$state_county <- paste(svi$ST_ABBR, svi$county, sep="_")
```

Long-term care facilities
```{r}
#Assume that if not listed, 0 retirement care available in county
retire <- read.table("/Users/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/combine17.txt", sep=",", header=T)
#Subset only retire homes based on NAICS number
retire <- retire[retire$NAICS %in% c('6233', '623311', '623311', '623312'),]
#Get rid of the "state total" rows
retire <- retire[!(retire$CTYDSCR %in% grep('Total', retire$CTYDSCR, value = T)), ]
retire <- retire[retire$SDSCR != "United States",]
#Remove "parish" and capitalize city
retire$CTYDSCR <- gsub('\\sParish', '', retire$CTYDSCR)
retire$CTYDSCR <- gsub('city', 'City', retire$CTYDSCR)
#Remove city
retire$CTYDSCR <- gsub(' City', '', retire$CTYDSCR)
#Get state abbreviations in from a key file
state.abb <- read.csv("/Users/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/state_abb_key.csv")
# state.abb <- read.csv("/Volumes/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/state_abb_key.csv")
#Make a new column with state abbreviations
state.abb$State <- as.character(state.abb$State)
state.abb$State <- gsub('^\\s', '', state.abb$State)
state.abb$Abb <- as.character(state.abb$Abb)
retire$SDSCR <- as.character(retire$SDSCR)
#Loop through and grab the abbreviation for each state
retire$state_abb <- 1:nrow(retire)
for (i in 1:nrow(retire)) {
  retire$state_abb[i] <- (state.abb$Abb[state.abb$State == retire$SDSCR[i]])
}
retire$state_county <- paste(retire$state_abb, retire$CTYDSCR, sep="_")
#Grab the state_county values from svi df that are not in the retire df because those counties have 0 retirement communities, then bind them to the retire df
retire.out <- rbind(subset(retire, select=c("state_county", "EST")), data.frame(state_county=svi$state_county[!(svi$state_county %in% retire$state_county)], EST=as.factor(rep(0, length(svi$state_county[!(svi$state_county %in% retire$state_county)])))))
retire.out$EST <- as.integer(as.character(retire.out$EST))
# nrow(retire.out[retire.out$EST > 0, ]) #Perfect
#Give it a better name
colnames(retire.out)[colnames(retire.out) == "EST"] <- "retire.comm"
```


Get data together
It looks like the covid-19 dataframe has the least number of counties observed (it does have counties with 0, so it's not that the rest are 0), subset all other DFs to match
Area, pop, coverage, hosp, salary, pres
```{r}
#First, subset only the columns that we want from each DF
covid.out <- subset(covid, select = c("Confirmed", "Deaths", "state_county"))
area.out <- subset(area, select = c("LND110210D", "state_county")); colnames(area.out) <- c("area", "state_county")
# pop.out <- subset(pop, select = c("POPESTIMATE2019", "state_county")); colnames(pop.out) <- c("population", "state_county") #SVI contains more pop info
coverage.out <- subset(coverage, select = c("Uninsured...", "state_county")); colnames(coverage.out) <- c("uninsured", "state_county")
salary.out <- subset(salary, select = c("Annual.Average.Weekly.Wage", "state_county")); colnames(salary.out) <- c("weekly.wage", "state_county")
pres.out <- subset(pres, select = c("trump.per", "age65andolder_pct", "rural_pct", "ruralurban_cc", "lesscollege_pct", "state_county")); pres.out$trump.per <- pres.out$trump.per * 100
covid.test.out <- subset(covid.test, select=c("test.per.pop", "State"))
svi.out <- subset(svi, select = c("E_TOTPOP", "SPL_THEME1", "MP_CROWD", "state_county"))
colnames(svi.out) <- c("pop", "svi", "crowding", "state_county")

# nrow(covid[covid$state_county %in% grep("LA", covid$state_county, value=T), ])
# nrow(area[area$state_county %in% grep("LA", area$state_county, value=T), ])
# nrow(pop[pop$state_county %in% grep("LA", pop$state_county, value=T), ]) #Problem
# nrow(coverage[coverage$state_county %in% grep("LA", coverage$state_county, value=T), ])
# nrow(salary[salary$state_county %in% grep("LA", salary$state_county, value=T), ])
# nrow(pres[pres$state_county %in% grep("LA", pres$state_county, value=T), ])
# nrow(prev.2day[prev.2day$state_county %in% grep("LA", prev.2day$state_county, value=T), ])
# nrow(prev.day.merged.out[prev.day.merged.out$state_county %in% grep("LA", prev.day.merged.out$state_county, value=T), ])

#Merge them all together with nested merge calls
dat.final <- merge(
  merge(
    merge(
      merge(
        merge(
          merge(
            merge(
              merge(
                covid.out,
                prev.day.merged.out, by = "state_county"),
                area.out, by = "state_county"),
                svi.out, by = "state_county"), #Replaces pop out b/c has better pop data
                coverage.out, by="state_county"),
                hosp.count.df, by="state_county"),
                salary.out, by="state_county"),
                pres.out, by="state_county"),
                retire.out, by="state_county")
dat.final$pop.dens <- dat.final$pop/dat.final$area
#Repeat tests per pop for each state
#Make a state column
dat.final$state <- as.character(gsub("\\_[A-Za-z \\,\\.\\s\\-\\'\\-]+", "", dat.final$state_county))
dat.final$tests.per.pop <- rep(1, nrow(dat.final))
for (i in 1:nrow(dat.final)) {
  dat.final$tests.per.pop[i] <- covid.test.out$test.per.pop[covid.test.out$State == dat.final$state[i]]
}
#Remove NAs
dat.final <- dat.final[!(is.na(dat.final$prev.ratio)), ]
#Make pop density into log pop density
dat.final$log.pop.dens <- log(dat.final$pop.dens)
dat.final$log.pop <- log(dat.final$pop)
#Remove Bedford County, VA b/c two of them
dat.final <- dat.final[dat.final$state_county != "VA_Bedford", ]
#Verify format
dat.final$uninsured <- as.numeric(as.character(dat.final$uninsured))
dat.final$weekly.wage <- as.numeric(as.character(gsub(',', '', dat.final$weekly.wage)))
dat.final$ruralurban_cc <- as.factor(dat.final$ruralurban_cc)
# str(dat.final)
```

Center and scale data
```{r}
dat.final[,c(5:6,8:15,17:19,21:23)] <- scale(dat.final[,c(5:6,8:15,17:19,21:23)], center = TRUE, scale = T)
# str(dat.final)
```


Use basis functions for predictors
```{r}
library(splines)
#Make the basis functions with 10 df
k <- 5
b.log.pop.dens <- bs(dat.final$log.pop.dens, df=k, intercept=T)
b.prev.diff <- bs(dat.final$prev.diff, df=k, intercept=T)
b.trump.per <- bs(dat.final$trump.per, df=k, intercept=T)
b.uninsured <- bs(dat.final$uninsured, df=k, intercept=T)
b.hosp <- bs(dat.final$hosp, df=k, intercept=T)
b.weekly.wage <- bs(dat.final$weekly.wage, df=k, intercept=T)
b.age56andolder_pct <- bs(dat.final$age65andolder_pct, df=k, intercept=T)
b.lesscollege_pct <- bs(dat.final$lesscollege_pct, df=k, intercept=T)
b.rural_pct <- bs(dat.final$rural_pct, df=k, intercept=T)
b.tests.per.pop <- bs(dat.final$tests.per.pop, df=k, intercept=T)
b.svi <- bs(dat.final$svi, df=k, intercept=T)
b.crowding <- bs(dat.final$crowding, df=k, intercept=T)
b.retire.comm <- bs(dat.final$retire.comm, df=k, intercept=T)
b.log.pop <- bs(dat.final$log.pop, df=k, intercept=T)

#Create X matrix to model true infection rate
#Create tensor product splines by pairwise multiplying the univariate splines
X <- NULL
for (i in 1:ncol(b.log.pop.dens)) {
  for (j in 1:ncol(b.prev.diff)) {
    # for (k in 1:ncol(b.rural_pct)) {
      for (l in 1:ncol(b.svi)) {
          # for (m in 1:ncol(b.crowding)) {
            for (n in 1:ncol(b.retire.comm)) {
              X <- cbind(X, b.log.pop.dens[,i]*b.prev.diff[,j]*b.svi[,l]*b.retire.comm[,n])
            }
          # }
      }
    # }
  }
}
X    <- as.matrix(X[,apply(X,2,max)>0.1])  # Remove basis function that are near zero for all sites
X    <- ifelse(X>0.001,X,0)
X <- cbind(rep(1,nrow(X)), X)
p <- ncol(X)

#Create Z matrix to model "underestimation proportion"
#Create tensor product splines by pairwise multiplying the univariate splines
Z <- NULL
# for (i in 1:ncol(b.trump.per)) {
  for (j in 1:ncol(b.uninsured)) {
    for (k in 1:ncol(b.hosp)) {
      for (l in 1:ncol(b.tests.per.pop)) {
          # for (m in 1:ncol(b.log.pop)) {
    Z <- cbind(Z, b.uninsured[,j]*b.tests.per.pop[,l])
          # }
      }
    }
  }
# }

Z    <- Z[,apply(Z,2,max)>0.1]  # Remove basis function that are near zero for all sites
Z    <- ifelse(Z>0.001,Z,0)
#Add intercept
Z <- cbind(rep(1,nrow(Z)), Z)
pz <- ncol(Z)
```

## Model with theta predicted

Calculations for informative prior for gamma[1] based on antibody testing:

NY city pop: 8398748; antibody prop: 0.21; antibody est: 1763737; confirmed: 141235; underestimation prop: 0.0801
Westchester county pop: 967506; antibody prop: 0.12; antibody est: 116101; confirmed: 24306; underestimation prop: 0.2093
Rockland county pop: 325789; antibody prop: 0.12; antibody est: 39095; confirmed: 9457; underestimation prop: 0.2419

LA county pop: 10039107; antibody prop: 0.05; antibody est: 501955; confirmed: 13823; underestimation prop: 0.0275

Note how much lower (worse) the underestimation proportion is for LA county. But we've also heard it was biased towards testing people who were more likely to be infected than expected at random, so it would be biased towards having a higher antibody prop. Let's assume it was actually 0.03 instead.

LA county pop: 10039107; antibody prop: 0.03; antibody est: 301173; confirmed: 13823; underestimation prop: 0.04589

It's still really low. It might also be that NY has been able to do regular testing more?

Based on all of this, let's set the prior to have 95% of its density on [0.03,0.25], which in logit scale is [-3.4761,-1.0986]. So the mean is the average of -2.287355 and the sd is 2.377487/2/1.96=0.6065018. Thus, we want a Normal(-2.287355,0.3678444).

For the informative prior for beta[1], use similar logic to get 95% of density on [0.02,0.2], so we want mean -2.63906, var=0.3923538

Need to transform the JAGS model into the Negative Binomial parameterization that STAN uses
```{r}
stan.nb4.stancode <- '
data {
  int<lower=1> N;                 //Number of observations
  int<lower=1> K;                 //Number of predictors in X
  int<lower=1> GK;                //Number of predictors in Z
  int<lower=1> pop[N];              //Population value
  matrix[N,K] X;                  //Fixed effects model matrix
  matrix[N,GK] Z;                 //Fixed effects model matrix for theta
  int y[N];                      //Response variable
}

//transformed data{
//  real logpop[N];
  
//  logpop = log(pop);
//}

parameters {
  real<lower=0,upper=100> r; //Size parameter for NB
  vector<lower=0.001,upper=1>[N] theta; //Underestimation proportion parameter
  vector<lower=0.001,upper=1>[N] lambda; //True infection rate
  real<lower=0> rr; //Sample size parameter for beta regression of theta
  real<lower=0> lr; //Sample size parameter for beta regression of lambda
  //real mu_b; //Hypermean parameter for betas
  real<lower=0> tau_b; //Hypervariance parameter for betas
  real<lower=0> tau_g; //Hypervariance parameter for gammas
  vector[K] beta_raw; //X regression params
  vector[GK] gamma_raw; //Z regression params
}

transformed parameters {
  real<lower=0.0001> m[N]; //Mean of NB
  //vector<lower=0.0001>[N] lambda; //True infection rate
  //vector<lower=0.0001>[N] theta; //Underestimation proportion
  vector<lower=0,upper=1>[N] p; //Mean of beta distribution
  vector<lower=0,upper=1>[N] q; //Mean of beta distribution
  vector[K] beta;
  vector[GK] gamma;

  beta = beta_raw * tau_b;
  gamma = gamma_raw * tau_g;
  q = inv_logit(X*beta); //lambda is exp of X*B
  p = inv_logit(Z*gamma); //p is inverse-logit of Z*G
  for (i in 1:N) {
    m[i] = pop[i]*theta[i]*lambda[i]; //Mean of NB is population x theta x true.rate (lambda)
  }

}

model {
  //mu_b ~ normal(0, 1);
  tau_b ~ cauchy(0, 2.5);
  tau_g ~ cauchy(0, 2.5);
  //to_vector(lambda) ~ beta(1,1);
  //target += lambda - 2 * log(1 + exp(lambda)); 
  //theta ~ beta(31, 835); //Informative prior from LA data
  r ~ cauchy(0, 2.5); //Prior on size parameter for NB
  for (k in 2:K) {
    beta_raw[k] ~ double_exponential(0, 1); //Normal shrinkage prior for betas with non-zero mean
  }
  beta_raw[1] ~ normal(-2.639057, 0.3923538);
  gamma_raw[1] ~ normal(-2.287355,0.3678444); //Informative prior for intercept from NY antibody testing (see note above)
  for (g in 2:GK) {
    gamma_raw[g] ~ double_exponential(0, 1); //LASSO shrinkage prior for gammas
  }
  
  //Likelihood
  for (j in 1:N) {
    lambda[j] ~ beta(lr*q[j], lr*(1-q[j]));
    theta[j] ~ beta(rr*p[j],rr*(1-p[j]));
    y[j] ~ neg_binomial(r, r/m[j]); //Stans beta is r/m
  }
}

generated quantities {
  int y_rep[N];                           //Vector for posterior predictive distribution
  vector[N] log_lik;                            //Vector for log-likelihood values for WAIC calculation
  vector[N] true_inf;
  
  //log_lik = neg_binomial_2_log_glm_lpmf(y | X, alpha, beta, phi);
  //y_rep = neg_binomial(r, r/m);
  
  for (a in 1:N) {
    y_rep[a] = neg_binomial_rng(r, r/m[a]);
    log_lik[a] = neg_binomial_lpmf(y[a] | r, r/m[a]);
    true_inf[a] = lambda[a]*pop[a];
  }

}

'
```

Define the data for STAN
```{r}
stan.nb4.standata <- list(
  y=dat.final$Confirmed, 
  N=nrow(X),
  X=X,
  Z=Z,
  pop=dat.final$pop,
  K=ncol(X),
  GK=ncol(Z)
)
```

Define initial values
```{r}
initf2 <- function(chain_id = 1) {
  # cat("chain_id =", chain_id, "\n")
  list(beta = array(c(0.05, rep(0, ncol(X)-1)), dim=ncol(X)),#rnorm(ncol(X), sd=0.1), dim = c(ncol(X))),
       beta_raw = array(c(0.05, rep(0, ncol(X)-1)), dim=ncol(X)),
       gamma = array(c(0.03, rep(0, ncol(Z)-1)), dim=ncol(Z)),
       gamma_raw = array(c(0.03, rep(0, ncol(Z)-1)), dim=ncol(Z)),
       theta = array(rep(0.01, nrow(X)), dim=nrow(X)),
       r=1, #Size parameter for NB
       rr=1, #Variance for beta likelihood
       lr=1,
       # mu_beta=0, #Hypermean for beta parameters
       tau_b=1, #Var of Cauchy for betas
       tau_g=1, #Var of Cauchy for gammas
       lambda=array(rep(0.01,nrow(X)),dim=nrow(X)),
       m=array(rep(0.01,nrow(X)),dim=nrow(X))
       )
}

# generate a list of lists to specify initial values
n_chains <- 4
init_ll <- lapply(1:n_chains, function(id) initf2(chain_id = id))
```


Run the model
```{r}
stan.nb4.stanfit <- stan(
  model_code = stan.nb4.stancode,
  data=stan.nb4.standata, 
  control = list(max_treedepth = 15, adapt_delta=0.9),
  init = init_ll,
  chains=4,
  warmup=300,
  iter=600,
  cores=4
)

traceplot(stan.nb4.stanfit, pars = c("r", "rr", "tau_b", "tau_g"))
traceplot(stan.nb4.stanfit, pars = c("lambda[51]"))
print(stan.nb4.stanfit, pars = "lambda")
print(stan.nb4.stanfit, pars=c("theta"))
stan.nb4.est.true <- as.data.frame(stan.nb4.stanfit, pars="true_inf")
stan.nb4.y_rep <- as.data.frame(stan.nb4.stanfit, pars="y_rep")
dat.final$stan.nb4.est.true <- apply(stan.nb4.est.true, 2, mean)
dat.final$stan.nb4.yrep <- apply(stan.nb4.y_rep, 2, mean)
#Look at predicted vs. true values
qplot(log(Confirmed), log(stan.nb4.yrep), data=dat.final)+geom_smooth(method="lm")
# summary(lm(log(stan.nb4.yrep) ~ log(Confirmed), data=dat.final))
#Calculate Gelman's Bayesian R2
dat.final$stan.nb4.resids <- dat.final$stan.nb4.yrep - dat.final$Confirmed
var(dat.final$stan.nb4.yrep)/(var(dat.final$stan.nb4.yrep)+var(dat.final$stan.nb4.resids)) #Bayesian r-squared of 0.684
```

There were 2 chains where the estimated Bayesian Fraction of Missing Information was low. See
http://mc-stan.org/misc/warnings.html#bfmi-lowExamine the pairs() plot to diagnose sampling problems
The largest R-hat is 1.69, indicating chains have not mixed.
Running the chains for more iterations may help. See
http://mc-stan.org/misc/warnings.html#r-hatBulk Effective Samples Size (ESS) is too low, indicating posterior means and medians may be unreliable.
Running the chains for more iterations may help. See
http://mc-stan.org/misc/warnings.html#bulk-essTail Effective Samples Size (ESS) is too low, indicating posterior variances and tail quantiles may be unreliable.
Running the chains for more iterations may help. See
http://mc-stan.org/misc/warnings.html#tail-ess

Make a map
```{r}
# install.packages(c("ggspatial", "libwgeom", "sf", "rnaturalearth", "rnaturalearthdata", "googleway"))
# install.packages(c("rgeos", "maps", "lwgeom", "ggrepel"))
library(ggplot2); library(scales); library(ggstance)
library(ggspatial); library(sf); library(lwgeom); library(ggrepel)
library(rnaturalearth); library(rnaturalearthdata); library(googleway)
library(rgeos)
library(maps)
library(gdata)

theme_set(theme_bw())

#Make base map
world <- ne_countries(scale="medium", returnclass = "sf")
states <- st_as_sf(map("state", plot = FALSE, fill = TRUE))
states <- cbind(states, st_coordinates(st_centroid(states)))
counties <- st_as_sf(map("county", plot = FALSE, fill = TRUE))
#Get state column
counties$state <- gsub("\\,[a-z \\,\\.\\s\\-\\'\\-]+", "", counties$ID)
#Turn first letter into uppercase
counties$state2 <- gsub('^([a-z])', '\\U\\1', gsub('\\s([a-z])', ' \\U\\1', counties$state, perl = T), perl = T)
counties$state2[counties$state2 == "District Of Columbia"] <- "District of Columbia"
#Get county column
counties$county <- gsub('^[A-Za-z ]+\\,', '', counties$ID)
#Turn first letter and letter after spaces into uppercase
counties$county <- gsub('^([a-z])', '\\U\\1', gsub('\\s([a-z])', ' \\U\\1', counties$county, perl = T), perl = T)
#Get state abbreviations in from a key file
state.abb <- read.csv("/Users/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/state_abb_key.csv")
# state.abb <- read.csv("/Volumes/tschappe/Documents/NC\ State\ Classes/ST540/Final\ Project/state_abb_key.csv")
#Make a new column with state abbreviations
state.abb$State <- as.character(state.abb$State)
state.abb$State <- gsub('^\\s', '', state.abb$State)
state.abb$Abb <- as.character(state.abb$Abb)
counties$state2 <- as.character(counties$state2)
#Loop through and grab the abbreviation for each state
counties$state_abb <- 1:nrow(counties)
for (i in 1:nrow(counties)) {
  counties$state_abb[i] <- (state.abb$Abb[state.abb$State == counties$state2[i]])
}
counties$state_county <- paste(counties$state_abb, counties$county, sep="_")
#Subset counties dataframe to match what dat.final has
counties2 <- counties[counties$state_county %in% dat.final$state_county, ]
#Loop through and grab the correct estimated infection values from final.dat. dat.final$state_county has punctuation, so remove it first
counties2$est.true.inf <- 1:nrow(counties2)
for (i in 1:nrow(counties2)) {
  counties2$est.true.inf[i] <- (dat.final$stan.nb4.est.true[gsub("[\\,\\.\\-\\'\\-]+", "", dat.final$state_county) == counties2$state_county[i]])
}
# counties <- subset(counties, grepl("north carolina", counties$ID))
counties2$est.true.inf <- as.numeric(counties2$est.true.inf)

(infection.map1 <- ggplot(data = world) +
  geom_sf() +
  geom_sf(data = counties2, aes(fill = log(counties2$est.true.inf)))+
  # geom_label_repel(data = sites.coords, mapping = aes(
  #   x = X,
  #   y = Y,
  #   label = ID,
  #   fill=Genotype),
  #   size=4
  # )+
  # scale_fill_manual(name="G143A Allele", breaks = c("H", "M", "W"), labels = c("Mixed", "Resistant", "Susceptible"), values = c(ggcols[4], ggcols[1], ggcols[3]))+
  coord_sf(xlim = c(-126, -66), ylim = c(24, 51), expand = FALSE))

(infection.map2 <- ggplot(data = world) +
  geom_sf() +
  geom_sf(data = counties2, aes(fill = log(counties2$est.true.inf)))+
  # geom_label_repel(data = sites.coords, mapping = aes(
  #   x = X,
  #   y = Y,
  #   label = ID,
  #   fill=Genotype),
  #   size=4
  # )+
  # scale_fill_manual(name="G143A Allele", breaks = c("H", "M", "W"), labels = c("Mixed", "Resistant", "Susceptible"), values = c(ggcols[4], ggcols[1], ggcols[3]))+
  coord_sf(xlim = c(-86, -68), ylim = c(24, 46), expand = FALSE))
```

```{r}
library(HDInterval)
hdis <- apply(stan.nb4.est.true, 2, FUN = function(x) hdi(x, credMass = 0.95))
hdis <- t(hdis)
# hdis.lambda <- apply(stan.nb4.lambda, 2, FUN = function(x) hdi(x, credMass = 0.95))
# hdis.lambda <- t(hdis.lambda)
hdis.all <- data.frame(
  state_county=dat.final$state_county[dat.final$state_county %in% c("NC_Wake", "NC_Mecklenburg", "IL_Cook", "CA_Los Angeles", "NY_Westchester", "AZ_Maricopa", "WA_King", "GA_Fulton")],
  inf.lower.HDI=hdis[dat.final$state_county %in% c("NC_Wake", "NC_Mecklenburg", "IL_Cook", "CA_Los Angeles", "NY_Westchester", "AZ_Maricopa", "WA_King", "GA_Fulton"),1],
  inf.upper.HDI=hdis[dat.final$state_county %in% c("NC_Wake", "NC_Mecklenburg", "IL_Cook", "CA_Los Angeles", "NY_Westchester", "AZ_Maricopa", "WA_King", "GA_Fulton"),2]
  # prop.lower.HDI=hdis.lambda[dat.final$state_county %in% c("NC_Wake", "NC_Mecklenburg", "IL_Cook", "CA_Los Angeles", "NY_Westchester", "AZ_Maricopa", "WA_King", "GA_Fulton"),1],
  # prop.upper.HDI=hdis.lambda[dat.final$state_county %in% c("NC_Wake", "NC_Mecklenburg", "IL_Cook", "CA_Los Angeles", "NY_Westchester", "AZ_Maricopa", "WA_King", "GA_Fulton"),2]
  )
```

